---
title: "Analyse Data for Covid Mexico"
output: html_document
---

Model 1 - Logistic Regression
Model 2 - Random Forest
Model 3 - KNN

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(haven)
library(dplyr)
library(corrplot)
library(ggplot2)
library(GGally)
library(ggfortify)
library(DBI)
library(tidyr)
library(factoextra)
library(cluster)
library(car)
library(olsrr)
library(Metrics)
library(caret)
library(tree)
library(randomForest)
library(gbm)
library(caTools)
library(ROCR)
library(glmnet)
library(rpart)
library(class)
library(rpart)
library(rpart.plot)
```


```{r}
setwd("~/Resume/Analyse Data for Covid Mexico")
```

```{r}
CovidData <- read.csv(file = '~/Resume/Analyse Data for Covid Mexico/Covid Data.csv')
```

Removing Variables (Due to computer specification)
```{r}
CovidData <- CovidData[-sample(1:nrow(CovidData), 948575),]
```



Checking 97/98/99 (Missing Values) inside the Database
```{r}
# Create an empty data frame with required columns
result_df <- data.frame(Column = character(),
                         Value_97 = numeric(),
                         Value_98 = numeric(),
                         Value_99 = numeric(),
                         stringsAsFactors = FALSE)

# Loop over each column in CovidData
for (col in colnames(CovidData)) {
  # Calculate the count of each value for the current column
  count_97 <- sum(CovidData[col] == 97)
  count_98 <- sum(CovidData[col] == 98)
  count_99 <- sum(CovidData[col] == 99)
  
  # Add a row to the data frame with the results for the current column
  result_df <- rbind(result_df, data.frame(Column = col,
                                           Value_97 = count_97,
                                           Value_98 = count_98,
                                           Value_99 = count_99,
                                           stringsAsFactors = FALSE))
}

# Print the resulting data frame
print(result_df)
```
**Comment: Pregnat/Intubed/ICU has a significant amount of 97/98/99 missing values, cannot simply remove them** 

Looking at individual column and the total as well as the type of data they contain

```{r}
aggregate(SEX ~ PREGNANT + SEX, data=CovidData, FUN=length)
```

Any values that is 97/98 will become 2 since PREGNAT == 2 is Male thus it does not make sense if they are pregnat
```{r}
CovidData$PREGNANT <- ifelse(CovidData$PREGNANT== 97, 2, CovidData$PREGNANT)
CovidData$PREGNANT <- ifelse(CovidData$PREGNANT== 98, 2, CovidData$PREGNANT)
```

Changing this set of code that contains "9999-99-99" to "No" and "Yes"
```{r}
CovidData$DEATH <- ifelse(CovidData$DATE_DIED == "9999-99-99", "No", "Yes")
```



```{r}
aggregate(PATIENT_TYPE ~ ICU + PATIENT_TYPE, data=CovidData, FUN=length)
```

Removing the columns as it contains too much missing data
```{r}
CovidData <- CovidData[, -which(names(CovidData) %in% c("INTUBED","ICU","DATE_DIED"))]
```


Removing all the remaining rows that contains 97/98/99
```{r}
for (col in colnames(CovidData)) {
  CovidData <- CovidData[!(CovidData[,col] %in% c(97, 98, 99)), ]
}
```


```{r}
CovidData$AgeGroup <- NA
CovidData[CovidData$AGE <= 20, "AgeGroup"] <- "0-20"
CovidData[CovidData$AGE > 20 & CovidData$AGE <= 40, "AgeGroup"] <- "21-40"
CovidData[CovidData$AGE > 40 & CovidData$AGE <= 59, "AgeGroup"] <- "40-59"
CovidData[CovidData$AGE >= 60, "AgeGroup"] <- "60++"
```

** In the event the codes do not work, we have to group 7 of the classification result together into just 2 group "1", "2" **
```{r}
CovidData$NEW_CLASIFFICATION_FINAL <- NA
CovidData[CovidData$CLASIFFICATION_FINAL <= 4, "NEW_CLASIFFICATION_FINAL"] <- "1"
CovidData[CovidData$CLASIFFICATION_FINAL > 4 , "NEW_CLASIFFICATION_FINAL"] <- "2"
```

Removing all the rows that contain 2 in their new classification status
```{r}
CovidData <- CovidData[CovidData$NEW_CLASIFFICATION_FINAL != 2,]
```

```{r}
CovidData <- CovidData[, -which(names(CovidData) %in% c("AGE","CLASIFFICATION_FINAL", "MEDICAL_UNIT", "NEW_CLASIFFICATION_FINAL"))]
```

Changing all the variables into factor to apply classification techniques
```{r}
CovidData$USMER <- as.factor(CovidData$USMER)
CovidData$SEX <- as.factor(CovidData$SEX)
CovidData$PATIENT_TYPE <- as.factor(CovidData$PATIENT_TYPE)
CovidData$PNEUMONIA <- as.factor(CovidData$PNEUMONIA)
CovidData$PREGNANT <- as.factor(CovidData$PREGNANT)
CovidData$DIABETES <- as.factor(CovidData$DIABETES)
CovidData$COPD <- as.factor(CovidData$COPD)
CovidData$ASTHMA <- as.factor(CovidData$ASTHMA)
CovidData$INMSUPR <- as.factor(CovidData$INMSUPR)
CovidData$HIPERTENSION <- as.factor(CovidData$HIPERTENSION)
CovidData$OTHER_DISEASE <- as.factor(CovidData$OTHER_DISEASE)
CovidData$CARDIOVASCULAR <- as.factor(CovidData$CARDIOVASCULAR)
CovidData$OBESITY <- as.factor(CovidData$OBESITY)
CovidData$RENAL_CHRONIC <- as.factor(CovidData$RENAL_CHRONIC)
CovidData$TOBACCO <- as.factor(CovidData$TOBACCO)
CovidData$DEATH <- as.factor(CovidData$DEATH)
CovidData$AgeGroup <- as.factor(CovidData$AgeGroup)
```


```{r}
death_freq <- table(CovidData$DEATH)

# Convert the result to a data frame
df_death_freq <- data.frame(DEATH = names(death_freq),
                            Frequency = as.numeric(death_freq))

# Create a bar plot using ggplot2
ggplot(df_death_freq, aes(x = DEATH, y = Frequency, fill = DEATH)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of DEATH variable",
       x = "DEATH", y = "Frequency") +
  scale_fill_manual(values = c("#00AFBB", "#FC4E07"), labels = c("No", "Yes"))
```
Comment: Too much "No" in "Death" column, might cause model to be bias


Random Under Sample
Note: Doing this might introduce “distortion” to the data values composition, hence giving a model which is not accurate
```{r}
count_df <- CovidData %>% count(DEATH)
n_no <- count_df$n[which(count_df$DEATH == "No")]
n_yes <- count_df$n[which(count_df$DEATH == "Yes")]
CovidData <- CovidData %>%
  group_by(DEATH) %>%
  slice_sample(n = min(n_no, n_yes), replace = FALSE)
```

```{r}
death_freq <- table(CovidData$DEATH)

# Convert the result to a data frame
df_death_freq <- data.frame(DEATH = names(death_freq),
                            Frequency = as.numeric(death_freq))

# Create a bar plot using ggplot2
ggplot(df_death_freq, aes(x = DEATH, y = Frequency, fill = DEATH)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency of DEATH variable",
       x = "DEATH", y = "Frequency") +
  scale_fill_manual(values = c("#00AFBB", "#FC4E07"), labels = c("No", "Yes"))
```


Logistic Regression 
```{r}
train <- sample.split(Y=CovidData$DEATH, SplitRatio=0.7)
trainset <- subset(CovidData, train==T)
testset <- subset(CovidData, train==F)
model <- glm(DEATH ~.,family=binomial(link='logit'),data = trainset)
summary(model)
```

```{r}
OR <- exp(coef(model))
OR
```

Checking accuracy/precision/recall/F1 score on training result
```{r}
prob.train <- predict(model, type = "response")
predict.y.train <- ifelse(prob.train > 0.5, "Yes", "No")
table.train <- table(trainset$DEATH, predict.y.train)
table.train
prop.table(table.train)
mean(predict.y.train == trainset$DEATH)
```

Checking accuracy/precision/recall/F1 score on test result
```{r}
prob.test <- predict(model, newdata = testset, type = "response")
predict.y.test <- ifelse(prob.test > 0.5, "Yes", "No")
table.test <- table(testset$DEATH, predict.y.test)

accuracy <- mean(predict.y.test == testset$DEATH)
precision <- table.test[2,2] / (table.test[2,2] + table.test[1,2])
recall <- table.test[2,2] / (table.test[2,2] + table.test[2,1])
f1_score <- 2 * precision * recall / (precision + recall)
table.test
cat("\n")
prop.table(table.test)
cat("\n")
cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision, "\n")
cat("Recall:", recall, "\n")
cat("F1 Score:", f1_score, "\n")
```

```{r}
pr.glm1 <- prediction(prob.test, testset$DEATH)
prf.glm1 <- performance(pr.glm1, measure="tpr", x.measure="fpr")
plot(prf.glm1)
abline(a=0, b=1)
```





Random Forest
```{r}
model.rf <- randomForest(DEATH~., data=trainset, importance=TRUE)
model.rf
```

```{r}
varImpPlot(model.rf)
```

```{r}
prob.train.rf <- predict(model.rf, newdata=testset)
error <- mean(prob.train.rf != testset$DEATH)
print(paste('Accuracy', 1-error))
```

```{r}
predictions <- predict(model.rf, newdata = testset)
Confusion.Matrix.RF <- confusionMatrix(predictions, testset$DEATH)
Confusion.Matrix.RF
```



K-Nearest Neighbors
```{r}
set.seed(150)
# train knn model
model_knn <- train(DEATH ~ ., data = trainset, method = "knn", trControl = trainControl(method = "cv", number = 10))
print(model_knn)
```

```{r}
# predicting on the test set and calculating the accuracy
predictions_knn <- predict(model_knn, newdata = testset)
accuracy_knn <- mean(predictions_knn == testset$DEATH)
#Generating the confusion matrix
conf_matrix_knn <- confusionMatrix(predictions_knn, testset$DEATH)
prec_knn <- conf_matrix_knn$byClass[1]
rec_knn <- conf_matrix_knn$byClass[2]
F1_knn <- (2 * prec_knn * rec_knn) / (prec_knn + rec_knn)

print(conf_matrix_knn$table)
cat("\n")
cat("Accuracy on testing data:", accuracy_knn, "\n")
cat("Precision:", prec_knn, "\n")
cat("Recall:", rec_knn, "\n")
cat("F1 Score:", F1_knn, "\n")
```
